Data Exploration: IUCN and AquaMaps

### 2015-09-04 data_explore.Rmd
fixed incorrect math for finding smaller range overlap with larger range

---
title: "Data Exploration: IUCN and AquaMaps"
author: "Jamie Afflerbach"
date: "9/2/2015"
output: html_document
---

```{r,echo=F,warning=FALSE,message=F}
library(ggplot2)
library(maps)
library(readr)
library(raster)
library(tidyr)
library(dplyr)

@@

```{r,echo=F,warning=FALSE,message=F,results='hide'}
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells.csv'), col_types = 'idiii', progress = TRUE)

@@

```{r,echo=F,eval=FALSE,warning=FALSE,message=F}
spp_list = spp_list%>%
              mutate(am_area = NA,
                     iucn_area=NA,
                     perc = NA,
                     am_perc=NA,
                     iucn_perc=NA)

for (i in 1:nrow(spp_list)){

@@

  spp <- spp_list[i,]$sciname

  spp_map <- get_spp_map(spp)

  am_spp_map = spp_map%>%
                dplyr::select(loiczid,am_sid)%>% #add in am_prob when we want to look at probability
                  unique()

#   r_am_spp  <-  subs(loiczid_raster, 
#                      am_spp_map[ , c('loiczid', 'am_sid')], 
#                      by = 'loiczid', 
#                      which = 'am_sid', 
#                      subsWithNA = TRUE)
#   
#   spp_list[i,]$am_area = area(r_am_spp,na.rm=T)%>%
#               cellStats(.,stat='sum')
#

  iucn_map <- spp_map %>%
    mutate(iucn_pres = iucn_area > 0)%>%
      dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
        unique()

@@

   if(TRUE %in% duplicated(iucn_map$loiczid))next()
#   
#   r_iucn_spp <- subs(loiczid_raster, 
#                      iucn_map[ , c('loiczid', 'iucn_pres')], 
#                     by = 'loiczid', 
#                      which = 'iucn_pres', 
#                      subsWithNA = TRUE)
#   
#   spp_list[i,]$iucn_area = area(r_iucn_spp,na.rm=T)%>%
#                 cellStats(.,stat='sum')

@@

cat(perc,perc_am,perc_iucn)

@@

>>>>

---
title: "Data Exploration: IUCN and AquaMaps"
output: html_document
---

# Librarires and Paths

library(ggplot2)
library(maps)
library(readr)
library(raster)
library(tidyr)
library(dplyr)

@@

```{r,echo=F,warning=FALSE,message=F,results='hide'}

# Data files

iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells.csv'), col_types = 'idiii', progress = TRUE)

@@

```{r,echo=F,eval=FALSE,warning=FALSE,message=F}

spp_list = spp_list%>%
              mutate(am_area = NA,
                     iucn_area=NA,
                     perc = NA,
                     am_perc=NA,
                     iucn_perc=NA,
                     sm_perc = NA, #filled with % of smaller range that is within the larger range
                     sm_range = NA) #what range is smaller

#there's likely a better way to do this than a for loop (lapply or sapply?)
for (i in 1:nrow(spp_list)){

@@

  spp <- spp_list[i,]$sciname #species name

  spp_map <- get_spp_map(spp) #get species map

  #get just the aquamaps cells
  am_spp_map = spp_map%>%
                  filter(!is.na(am_prob))%>%
                dplyr::select(loiczid,am_sid,am_prob)%>% #add in am_prob when we want to look at probability
                  unique()
  #rasterize aquamaps map
  r_am_spp  <-  subs(loiczid_raster, 
                     am_spp_map[ , c('loiczid', 'am_sid')], 
                     by = 'loiczid', 
                     which = 'am_sid', 
                     subsWithNA = TRUE)
#   
#   spp_list[i,]$am_area = area(r_am_spp,na.rm=T)%>%
#               cellStats(.,stat='sum')
#

  #get the iucn cells
  iucn_map <- spp_map %>%
    mutate(iucn_pres = iucn_area > 0)%>%
      dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
        unique()

@@

  #this is incorporated because there are some duplications for loiczids with IUCN cells. I'm guessing this is because of the parent/subpopulations?

   if(TRUE %in% duplicated(iucn_map$loiczid))next()
#   
#  rasterize IUCN map
  r_iucn_spp <- subs(loiczid_raster, 
                     iucn_map[ , c('loiczid', 'iucn_pres')], 
                    by = 'loiczid', 
                     which = 'iucn_pres', 
                     subsWithNA = TRUE)
#   
#   spp_list[i,]$iucn_area = area(r_iucn_spp,na.rm=T)%>%
#                 cellStats(.,stat='sum')

@@

  spp_list[i,]$sm_range = ifelse(length(cells_am)<length(cells_iucn),'AM','IUCN') #what dataset has the smaller range
  spp_list[i,]$sm_perc = length(cells_overlap)/min(length(cells_am),length(cells_iucn))*100 #this is the total percent of smaller range encompassed in the larger range
cat(perc,perc_am,perc_iucn)

@@

Looking at the percent of smaller species range that is within the larger range.

```{r,echo=F}

d <- spp_list%>%
  dplyr::select(sciname,sm_perc,sm_range)%>%
      unique()

ggplot(d,aes(sm_perc,fill=sm_range))+geom_histogram(binwidth=1,alpha=.7)+
  scale_fill_discrete(name="Dataset with smaller range",labels=c("AquaMaps", "IUCN"))+labs(x='Percent of smalle range that falls within the larger range', y='Number of Species')
